Requirements and Current State of Knowledge Representation Systems for Service Robotics

### 2012-04-17 krs-survey.tex
ainly added section on horn clauses

@@

Most (but not all) of the systems we survey rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the inference problem (called
\emph{decidability}: is a given logical sentence true in my model?) in a
tractable manner.

@@

The family of \emph{description logics}~\cite{Baader2008} also play an
important role. It is a subset of the first-order logic, with some extensions
in second-order logic. Description logics are notable because most of them are
known to be decidable. In description logic, axioms are build from
\emph{concepts}, \emph{roles} (that are unary or binary predicates) and
\emph{individuals}. The W3C OWL-DL standard is a widely-used language to describe
domains with the description logic.

@@

Generally speaking, expressive power comes at the cost of ease of
\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
at section~\ref{sect|reasoning}.} tests, possibly leading to undecidable formal
systems, \ie systems where it is proven that a proposition can not be decided
to be true or false.

The relationship between expressive power and the complexity of reasoning problem that
follow has been partially formalized for Description Logics.
Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
for Description Logics that allows to conveniently explore this relationship
and indexes most of the litterature on that subject.

@@

@@

\paragraph{Rule Languages}
\item generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
        \footnotesize \concept{looksAt(?agt, ?obj)} $\land$
        \concept{pointsAt(?agt,?obj)} \par $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
        \normalsize

@@

$Egg(a) \wedge Egg(b) \wedge MakeOmelette(a, b, c) \to \lnot Egg(a) \wedge
\lnot Egg(b) \wedge Omelette(c)$

The insertion of the proposition $MakeOmelette(a, b, c)$ leads to retraction of
other facts. This rule requires non-monotonic reasoning to be applied.

@@

Some systems \cite{Kunze??} rely on external dedicated physics engine to
compute symbolic facts from on-demand physics simulation. This, however,
requires a tight integration between the symbolic model and a geometric model
that carries the geometries and physical properties of objects.

@@

Memory has been studied at lenght in the cognitive neuroscience community, and
the ability to forget is proven to play a key role for the cognition.
\fxwarning{Add references -> ask Claire or Gladys}.

Most knowledge representation systems offers some kind of working memory as a
pool of facts that are not forgotten by the robot until it is halted. Some
systems may propose persistent storages that allow the robot knowledge to grow
over time, while other may offer a larger range of memory categories, like
short term memory (that last for a couple of seconds) or episodic memory (that
allows the robot to selectively remember facts associated to specific events).
\fxwarning{Add references of paper in the robotics community that have
explicitely tackled the memory issue}.

@@

>>>>

\subsection*{A Note on Terminology}

Since this article discusses at length the concept of knowledge in the context
of robotics, it is useful to make our terminology explicit.  No general
agreement on a definition of ``knowledge'' exists. In our context, we call
``knowledge'' \emph{an explicit set of interrelated logical facts that are
meaningful to the robot executive controller} (by \emph{meaningful} we mean
that can possibly be interpreted to lead to a purposeful action).

The relation of \emph{data} and \emph{information} to knowledge is a debated
epistemology question (the interested reader is invited to refer to the
Wikipedia page on the ``DIKW'' hierarchy). In this article, we will associate
data to low-level material like raw sensor output, and information to
uncontextualized symbolic facts.

To give a example, image a human watching at a book and being tracked by a
Kinect sensor: the pose of the human skeleton in the world would be the data,
the fact \concept{looksAt(human, book)} as computed by a geometric reasoning
module would be the information, the fact \concept{looksAt(john,
war\_and\_peace)}, fully grounded and connected to the whole knowledge base of
the robot would be proper knowledge.

@@

Most (but not all) of the systems we survey rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the inference problem (called
\emph{satisfiability}: is a given logical sentence true in my model?) in a
tractable manner.

@@

\emph{Horn clauses} are an important subset of FOL because the satisfiability
of a set of such clauses is a $P$-complete problem. A Horn clause is a
disjunction of literals (a \emph{clause}) with at most one positive literal:
$\neg p \vee \neg q \vee \cdots \vee \neg t \vee u$, which can also be
represented as $(p \wedge q \wedge \cdots \wedge t) \rightarrow u$.  Important
logic programming languages like Prolog are based on Horn clauses.

The family of \emph{Description Logics}~\cite{Baader2008} also play an
important role. It is also a subset of the first-order logic, with some
extensions in second-order logic. Description logics are notable because most
of them are known to be decidable. In description logic, axioms are build from
\emph{concepts}, \emph{roles} (that are unary or binary predicates) and
\emph{individuals}. The W3C OWL-DL standard is a widely-used language to
describe domains with the description logic.

@@

Generally speaking, expressive power comes at the cost of more complex
\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
at section~\ref{sect|reasoning}.} computations, possibly leading to
untractable, if not undecidable (\ie systems where it is proven that a
proposition can not be decided to be true or false) problems.

\begin{figure}
    \centering
    \includegraphics[width=0.6\columnwidth]{expressive_overlap_dl_horn.pdf}
    \caption{Expresiveness overlap of Description Logics and logic programs
    based on Horn clauses, taken from~\cite{Grosof2003}}
    \label{fig|overlap_dl_horn}
\end{figure}

Figure~\ref{fig|overlap_dl_horn} shows that the expresive power of description
logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
mention extensions to description logics based on rule systems to bring closer
the two approaches.

The relationships between expressive power and reasoning complexity that follow
has been extensively studied for Description Logics.
Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
that allows to conveniently explore these relationships and indexes most of the
litterature on that subject.

@@

Memory has been studied at lenght in the cognitive psychology and
neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
introduce the idea of \emph{short-term} and \emph{long-term} memory,
Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organizes
the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
into a hierarchy. Short-term memory is refined with the concept of
\emph{working memory} by Baddeley~\cite{Baddeley2010}
(Figure~\ref{fig|memory_models}).

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
    \caption{Overview of the main types of memories}
    \label{fig|memory_models}
\end{figure}

Most knowledge representation systems offers some kind of memory as a pool of
facts that are not forgotten by the robot until it is halted (this memory is
often refered as a \emph{working memory}, but with a meaning unrelated to
Baddeley's definition). Some systems may propose persistent storages that allow
the robot knowledge to grow over time, while other may offer a larger range of
memory categories, like short term memory (that lasts for a couple of seconds)
or episodic memory (that allows the robot to selectively remember facts
associated to specific events).

@@

\paragraph{Rule Languages}

As mentionned earlier, knowledge models based on description logics can be extended through rule languages.

Intersection of properties is an example of expression that can only be represented with rules: for instance,

generic SWRL ({\em Semantic Web Rule Language}) rules like: \par
        \footnotesize \concept{looksAt(?agt, ?obj)} $\land$
        \concept{pointsAt(?agt,?obj)} \par $\Rightarrow$ \concept{focusesOn(?agt, ?obj)}
        \normalsize

@@

\concept{Egg}$(a) \wedge $ \concept{Egg} $(b) \wedge $
\concept{MakeOmelette}$(a, b, c) \to \lnot $ \concept{Egg}$(a) \wedge \lnot $
\concept{Egg}$(b) \wedge $ \concept{Omelette}$(c)$

The insertion of the proposition \concept{MakeOmelette}$(a, b, c)$ leads to
retraction of other facts. This rule requires non-monotonic reasoning to be
applied.

@@

Some systems \cite{Kunze2011a} rely on external dedicated physics engine to
compute symbolic facts from on-demand physics simulation. This, however,
requires a tight integration between the symbolic model and a geometric model
that carries the geometries and physical properties of objects.

@@

@@

In the survey inclusion criteria, we have insisted on only considering robotic
systems that acquire knowledge by themselves, during their runtime.

\emph{Acquiring knowledge} means in our context building new statements
(usually as new logical facts) connected to our existing knowledge from
external sources of information. We consider mainly three of them:
proprioceptive/exteroceptive sensors, interaction with other agents, humans or
robots, and remote databases. This process has generally at least two steps:
the information acquisition by itself, and the \emph{transformation} of the
information into knowledge, \emph{aligned} with the robot existing model
(following our terminology for information and knowledge, as discussed in the
introduction).

Knowledge acquision is generally not done directly in the knowledge
representation system. On the contrary, several, if not numerous, external
components are usually required to convert percepts into symbolic facts and to
ground them.

While we do not review in this article all these systems, the whole process of
knowledge acquisition is central in cognitive robotic architecture and the
design of knowledge representation systems can influence or be influenced by
the approach to knowledge acquisition.

In particular, complex robotic systems often require multi-modal perception
capabilities (for instance, a robot can only interpret an utterance like ``this
is a plate'' if it is able to understand gestures, understand natural language
and merge them in a timely manner). Multi-modal interpretation can take place
at various levels, but in many cases (especially if the modalities are of very
different natures, like in the example above) merging will require
symbolic-level reasoning. The KRS has a direct impact on the feasability and
ease of such operations.
